---
title: "Critical Thinking Group 4 - HW2"
author: "Sreejaya, Suman, Vuthy"
date: "September 27, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
```


 1. Download the classification output data set
```{r}
classification.orig  <- read.csv("https://raw.githubusercontent.com/Nguyver/DATA621-HW/master/HW2/classification-output-data.csv", header=TRUE, sep=",", stringsAsFactors = FALSE)
```

2. The data set has three key columns we will use:

- **class**: the actual class for the observations
- **score.class**: the predicted class for the observations (based on a threshold of 0.5)
- **scored.probability**: the predicted probability o=f success for the observation

Use the table() function to get the raw confusion matris for this scored dataset. Make sure you understand the output. In particular, do the rows represent the actual or predicted class? The Columns?  

In our case, the columns are the predictions and the rows are the actuals.  
```{r}
class.df <- classification.orig
class.df$class <- factor(class.df$class)
class.df$scored.class <- factor(class.df$scored.class)

glimpse(class.df)
summary(class.df)
table(class.df$class, class.df$scored.class)

# Verifying Table results  
#               = Predict-Actual  
#True Positive  = 1-1  
#False Positive = 1-0  
#True Negative  = 0-0  
#False Negative = 0-1  
sum(class.df$class == class.df$scored.class & class.df$class ==1) #TP (2,2)
sum(class.df$class != class.df$scored.class & class.df$class ==0) #FN (1,2)

sum(class.df$class == class.df$scored.class & class.df$class ==0) #TN (1,1)
sum(class.df$class != class.df$scored.class & class.df$class ==1) #FP (2,1)
```

3. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the accuracy of the predictions.
$$Accuracy = \frac{TP+TN}{TP+FP+TN+FN}$$
```{r}
accuracy <- function(data, actual.idx=9, predict.idx=10){
  confusion <- table(data[,actual.idx], data[,predict.idx])
  
  value <- (confusion[2,2]+confusion[1,1])/(confusion[2,2]+confusion[2,1]+confusion[1,1]+confusion[1,2])
  
  return(value)
}

#Example
accuracy(class.df,9,10)
```

4. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the classification error rate of the predictions.
$$Classification Error Rate = \frac{FP+FN}{TP+FP+TN+FN}$$
Verify that you get an accuracy and error rate that sums to one.

```{r}
classErrorRate <- function(data, actual.idx=9, predict.idx=10){
  confusion <- table(data[,actual.idx], data[,predict.idx])
  
  value <- (confusion[1,2]+confusion[2,1])/(confusion[2,2]+confusion[2,1]+confusion[1,1]+confusion[1,2])
  
  return(value)
}

classErrorRate(class.df)+accuracy(class.df)
```

5. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the precision of the predictions.
$$Precision = \frac{TP}{TP+FP}$$

```{r}
precision <- function(data, actual.idx=9, predict.idx=10){
  confusion <- table(data[,actual.idx], data[,predict.idx])
  
  value <- confusion[2,2]/(confusion[2,2]+confusion[1,2])
  
  return(value)
}

precision(class.df)
```

6. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the sensitivity of the predictions. Sensitivity is also known as recall.
$$Sensitivity = \frac{TP}{TP+FN}$$
```{r}
sensitivity <- function(data, actual.idx=9, predict.idx=10){
  confusion <- table(data[,actual.idx], data[,predict.idx])
  
  value <- confusion[2,2]/(confusion[2,2]+confusion[2,1])
  
  return(value)
}

sensitivity(class.df)
```

7. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the specificity of the predictions.
$$Specificity = \frac{TN}{TN+FP}$$
```{r}
specificity <- function(data, actual.idx=9, predict.idx=10){
  
  
  confusion <- table(data[,actual.idx], data[,predict.idx])
  
  value <- confusion[1,1]/(confusion[1,1]+confusion[1,2])
  
  return(value)
}

specificity(class.df)
```

8. Write a function that takes the data set as a dataframe, with actual and predicted classifications identified, and returns the F1 score of the predictions.
$$F1 Score = \frac{2 X Precision X Sensitivity}{Precision + Sensitivity}$$
```{r}
f1Score <- function(data, actual.idx=9, predict.idx=10){
  confusion <- table(data[,actual.idx], data[,predict.idx])
  
  value <- (2*precision(data)*sensitivity(data))/(precision(data)+sensitivity(data))
  
  return(value)
}

f1Score(class.df)
```

9. Before we move on, let's consider a question that was asked: What are the bounds on the F1 score? Show that the F1 score will always be between 0 and 1. (Hint: If 0 < a < 1 and 0 < b < 1 then ab < a.)
```{r}

```

10. Write a function that generates an ROC curve from a data set with a true classification column (class in our example) and a probability column (scored.probability in our example). Your function should return a list that  includes  the  plot  of  the  ROC  curve  and  a  vector  that  contains  the  calculated  area  under  the  curve (AUC). Note that I recommend using a sequence of thresholds ranging from 0 to 1 at 0.01 intervals.

```{r}
library(ggplot2)

getRoCData <- function(data)
{
  rocResults <- data.frame(Threshold=c(), Sn=c(), Sp=c())
  
  thresholds <- seq(0, 1, by=0.01)
  for(threshold in thresholds)
  {
    data <- data %>% mutate(predicted = as.numeric(scored.probability >= threshold))
    data$predicted  <- factor( data$predicted , levels=c(0,1) )
    
    Sn <- sensitivity(data, actual.idx =  9, predict.idx =  12)
    Sp <- specificity(data, actual.idx =  9, predict.idx =  12)
    
    rocResults <- rbind(rocResults, data.frame(Threshold=c(threshold), Sn=c(Sn), Sp=c(Sp)))
  }
  
  rocPlot <- ggplot(rocResults, aes(x=1-Sp, y=Sn) ) + geom_line(size = 1, alpha = 0.6) +
    geom_point(data=rocResults[rocResults$Threshold == 0.5,], 
               aes(x=1-Sp, y=Sn), colour="orange") +
    geom_abline(slope=1, intercept=0, colour="blue", size=0.50) +
          xlim(0, 1) + ylim(0, 1) +
    labs(title="ROC Curve",
         x = "False Positive Rate ( 1 - Sp) ",
         y = "True Positive Rate (Sn)")

  #TODO - How to calculate the area under the curve ?
  auc <- c()
  results <- list(rocPlot, auc)

  return(results)
}

getRoCData(class.df)
```

11. Use your **created R function** and the provided classifcation output data set to produce all of the classification metrics discussed above.
```{r}

```

12. Investigate the **caret** package. In particular, consider the functions confusionMatrix, sensitivity, and specificity. Apply the functions to the data set. How do the results compare with your own functions?
```{r}

```

13. Investate the **pROC** package. Use it to generate an ROC curve for the data set. How do the results compare with your own functions?

```{r}

```