---
title: "Critical Thinking Group 4 - HW3"
author: "Sreejaya, Suman, Vuthy"
date: "October 10, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=FALSE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
```

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(car)
library(recommenderlab)
library(knitr)
library(PerformanceAnalytics)
library(Amelia)
```

## Overview

Dataset  
[Crime - Training data](https://github.com/Nguyver/DATA621-HW/blob/master/HW3/crime-training-data.csv)  
[Crime - Evaluation Data](https://github.com/Nguyver/DATA621-HW/blob/master/HW3/crime-evaluation-data.csv)

## Data Exploration

There are 466 observations, and 14 variables in the given training dataset. The below is a glimpse of the data.
```{r}
crime.trn  <- read.csv("https://raw.githubusercontent.com/Nguyver/DATA621-HW/master/HW3/crime-training-data.csv", header=TRUE, sep=",", stringsAsFactors = FALSE)

crime.evl  <- read.csv("https://raw.githubusercontent.com/Nguyver/DATA621-HW/master/HW3/crime-evaluation-data.csv", header=TRUE, sep=",", stringsAsFactors = FALSE)

glimpse(crime.trn)

```

# A visual take on the missing values might be helpful: the Amelia package has a special plotting function missmap() that will plot your dataset and highlight missing values:
```{r}

missmap(crime.trn, main = "Missing values vs observed")
```
There are no missing values in the dataset.


```{r}
length(crime.trn$zn)
table(crime.trn$zn)
```
Out of 466 values 339 are zeros. So we would like to treat zn as binary, land size over 25,000 sq.ft as 1 and below 25,000 sq.ft as 0

```{r}
crime.trn$zn <- ifelse(crime.trn$zn>0, 1, 0)
table(crime.trn$zn)

```


Lets check the summary of the given dataset, as well as check for any NA values in the data set.
```{r}
summary(crime.trn)
```


There appears to be no missing values. Lets plot the correlation between the variables.

```{r, warning=FALSE, message=FALSE}
cor.matrix <- cor(crime.trn[,1:ncol(crime.trn)])
chart.Correlation(cor.matrix, histogram=TRUE, pch=19)
```

From the above correlation matrix , the target varaible seems to have correlation with 

  * zn    - proportion of residential land zoned for large lots 
  * indus - proportion of non-retail business acres per suburb 
  * nox   - nitrogen oxides concentration
  * age   - proportion of owner-occupied units built prior to 1940
  * dis   - weighted mean of distances to five Boston employment centers  
  * rad   - index of accessibility to radial highways 
  * tax   - full-value property-tax rate per $10,000 
  * lstat - lower status of the population 


## Data Preparation

Lets look at each of the predictor variable's data:

### *zn* - proportion of residential land zoned for large lots

```{r}
#proportion of residential land zoned for large lots

hist(crime.trn$zn, main="Histogram of zn", xlab="zn", border="blue", col="green", las=1)
table(crime.trn$zn)

#Lets look at the average response rate.
#gives %ges across rows , and rounded by 2 decimals.
kable(round(prop.table(table(crime.trn$zn, crime.trn$target), 1),2))
```
```{r}

```

From the above, it appears like majority of the neighborhoods have no residential land zoned for large lots. When we looked at the average response rates for the zn data, we have identified following categories:

| Target  |           |
|---------|-----------|
|0, 1     |    zn     |
|---------|-----------|
|0.37,0.63|  0        |
|0.38,0.62|  20       |
|0.78,0.22|  22       |
|1.0,0.00 | others.   |
|---------|-----------|

So, we left with these 4 categories. So, by definition, we need to make 3 dummy variables.

```{r, warning=FALSE, message=FALSE}
crime.trn.new <-  crime.trn %>%
                    mutate(zn1 = as.numeric(zn==0),
                           zn2 = as.numeric(zn==20),
                           zn3 = as.numeric(zn==22)) %>%
                    select(-zn)

#since we prepared the dummy variables needed, we do not need the zn any more here.
kable(head(crime.trn.new,15))
```

Similarly, let's proceed with others

### *indus* - proportion of non-retail business acres per suburb

```{r}
hist(crime.trn$indus, main="Histogram of Indus", xlab="indus", border="blue", col="green", las=1)
#round(prop.table(table(crime.trn.new$indus, crime.trn.new$target), 1),2)
```


|Indus | Target    |
|------|-----------|
|2.95  | 1.00| 0.00|
|3.24  | 1.00| 0.00|
|3.33  | 1.00| 0.00|
|3.37  | 1.00| 0.00|
|3.41  | 1.00| 0.00|
|3.44  | 1.00| 0.00|
|3.64  | 1.00| 0.00|
|3.75  | 1.00| 0.00|
|3.78  | 1.00| 0.00|
|4     | 1.00| 0.00|
|4.05  | 1.00| 0.00|
|4.15  | 1.00| 0.00|
|4.39  | 1.00| 0.00|
|4.49  | 1.00| 0.00|
|4.86  | 1.00| 0.00|
|4.93  | 1.00| 0.00|
|4.95  | 1.00| 0.00|
|5.13  | 1.00| 0.00|
|5.19  | 1.00| 0.00|
|5.32  | 1.00| 0.00|
|5.64  | 1.00| 0.00|
|5.96  | 1.00| 0.00|
|6.06  | 1.00| 0.00|
|6.07  | 1.00| 0.00|
|6.09  | 1.00| 0.00|
|10.01 | 1.00| 0.00|
|10.81 | 1.00| 0.00|
|11.93 | 1.00| 0.00|
|12.83 | 1.00| 0.00|
|13.89 | 1.00| 0.00|
|13.92 | 1.00| 0.00|
|15.04 | 1.00| 0.00|
|6.41  | 1.00| 0.00|
|6.91  | 1.00| 0.00|
|7.07  | 1.00| 0.00|
|7.87  | 1.00| 0.00|
|25.65 | 1.00| 0.00|
|27.74 | 1.00| 0.00|
|------|-----|-----|
|7.38  | 0.67| 0.33|
|------|-----|-----|
|9.69  | 0.71| 0.29|
|10.59 | 0.70| 0.30|
|------|-----|-----|
|5.86  | 0.78| 0.22|
|6.96  | 0.80| 0.20|
|------|-----|-----|
|8.56  | 0.91| 0.09|
|------|-----|-----|
|9.9   | 0.18| 0.82|
|------|-----|-----|
|21.89 | 0.07| 0.93|
|------|-----|-----|
|18.1  | 0.00| 1.00|
|19.58 | 0.00| 1.00|
|8.14  | 0.00| 1.00|
|3.97  | 0.00| 1.00|
|6.2   | 0.00| 1.00|
|------|-----|-----|

The distribution above appears some what weired, and we could not find a meaningful categorization here.

### *nox*   - nitrogen oxides concentration 

```{r}
hist(crime.trn$nox, main="Histogram of nox", xlab="nox", border="blue", col="green", las=1)
#kable(round(prop.table(table(crime.trn.new$nox, crime.trn.new$target), 1),2))
```

|nox    | Target    |
|-------|-----------|
|0.389  | 1.00| 0.00|
|0.392  | 1.00| 0.00|
|0.394  | 1.00| 0.00|
|0.398  | 1.00| 0.00|
|0.4    | 1.00| 0.00|
|0.401  | 1.00| 0.00|
|0.403  | 1.00| 0.00|
|0.404  | 1.00| 0.00|
|0.405  | 1.00| 0.00|
|0.409  | 1.00| 0.00|
|0.41   | 1.00| 0.00|
|0.411  | 1.00| 0.00|
|0.413  | 1.00| 0.00|
|0.415  | 1.00| 0.00|
|0.4161 | 1.00| 0.00|
|0.422  | 1.00| 0.00|
|0.426  | 1.00| 0.00|
|0.428  | 1.00| 0.00|
|0.429  | 1.00| 0.00|
|0.433  | 1.00| 0.00|
|0.437  | 1.00| 0.00|
|0.4379 | 1.00| 0.00|
|0.439  | 1.00| 0.00|
|0.442  | 1.00| 0.00|
|0.4429 | 1.00| 0.00|
|0.445  | 1.00| 0.00|
|0.447  | 1.00| 0.00|
|0.448  | 1.00| 0.00|
|0.449  | 1.00| 0.00|
|0.453  | 1.00| 0.00|
|0.458  | 1.00| 0.00|
|0.46   | 1.00| 0.00|
|0.469  | 1.00| 0.00|
|0.472  | 1.00| 0.00|
|0.484  | 1.00| 0.00|
|0.488  | 1.00| 0.00|
|0.499  | 1.00| 0.00|
|0.51   | 1.00| 0.00|
|0.515  | 1.00| 0.00|
|0.518  | 1.00| 0.00|
|0.524  | 1.00| 0.00|
|0.547  | 1.00| 0.00|
|0.55   | 1.00| 0.00|
|0.573  | 1.00| 0.00|
|0.581  | 1.00| 0.00|
|0.609  | 1.00| 0.00|
|-------------------|
|0.52   | 0.91| 0.09|
|-------------------|
|0.493  | 0.67| 0.33|
|0.585  | 0.71| 0.29|
|-------------------|
|0.431  | 0.78| 0.22|
|0.489  | 0.79| 0.21|
|-------------------|
|0.464  | 0.88| 0.12|
|-------------------|
|0.544  | 0.18| 0.82|
|-------------------|
|0.624  | 0.07| 0.93|
|0.538  | 0.05| 0.95|
|-------------------|
|0.504  | 0.00| 1.00|
|0.507  | 0.00| 1.00|
|0.532  | 0.00| 1.00|
|0.575  | 0.00| 1.00|
|0.58   | 0.00| 1.00|
|0.583  | 0.00| 1.00|
|0.584  | 0.00| 1.00|
|0.597  | 0.00| 1.00|
|0.605  | 0.00| 1.00|
|0.614  | 0.00| 1.00|
|0.631  | 0.00| 1.00|
|0.647  | 0.00| 1.00|
|0.655  | 0.00| 1.00|
|0.659  | 0.00| 1.00|
|0.668  | 0.00| 1.00|
|0.671  | 0.00| 1.00|
|0.679  | 0.00| 1.00|
|0.693  | 0.00| 1.00|
|0.7    | 0.00| 1.00|
|0.713  | 0.00| 1.00|
|0.718  | 0.00| 1.00|
|0.74   | 0.00| 1.00|
|0.77   | 0.00| 1.00|
|0.871  | 0.00| 1.00|
|-------|-----|-----|

There is no meaningful categorization can be concluded from the above. Let's proceed with other variables.

### *age*   - proportion of owner-occupied units built prior to 1940 

```{r}
hist(crime.trn$age, main="Histogram of age", xlab="age", border="blue", col="green", las=1)
hist(crime.trn.new$age, main="Histogram of age", xlab="age", border="blue", col="green", las=1)
#kable(round(prop.table(table(crime.trn.new$age, crime.trn.new$target), 1),2))
```

Looks like the buildings with age > 100 are mentioned as 100 in the above. We could not derive a specific categorization here, so, we leave the variable as is.


### *dis*   - weighted mean of distances to five Boston employment centers 

```{r}
hist(crime.trn$dis, main="Histogram of dis", xlab="dis", border="blue", col="green", las=1)
hist(crime.trn.new$dis, main="Histogram of dis", xlab="dis", border="blue", col="green", las=1)
#kable(round(prop.table(table(crime.trn.new$dis, crime.trn.new$target), 1),2))
```

### *rad*   - index of accessibility to radial highways 
```{r}
hist(crime.trn$rad, main="Histogram of rad", xlab="rad", border="blue", col="green", las=1)
hist(crime.trn.new$rad, main="Histogram of rad", xlab="rad", border="blue", col="green", las=1)
#kable(round(prop.table(table(crime.trn.new$rad, crime.trn.new$target), 1),2))
```

### *tax*   - full-value property-tax rate per $10,000 
```{r}
hist(crime.trn$tax, main="Histogram of tax", xlab="tax", border="blue", col="green", las=1)
hist(crime.trn.new$tax, main="Histogram of tax", xlab="tax", border="blue", col="green", las=1)
#kable(round(prop.table(table(crime.trn.new$tax, crime.trn.new$target), 1),2))
```

### *lstat*  - lower status of the population  
```{r}
hist(crime.trn$lstat, main="Histogram of lstat", xlab="lstat", border="blue", col="green", las=1)
hist(crime.trn.new$lstat, main="Histogram of lstat", xlab="lstat", border="blue", col="green", las=1)
```  

Let's remove the predictors with low correlation with the target:

```{r}
crime.trn.new <-  crime.trn %>%
                    select(-chas, -rm, -ptratio, -black, -medv)
#crime.trn.new <-  crime.trn.new %>%
#                    select(-chas, -rm, -ptratio, -black, -medv)
```


Also, let's split our dataset into training (80%) and test (20%).

```{r}
set.seed(999) 
s=sample(1:nrow(crime.trn.new),0.80*nrow(crime.trn.new)) 
crime.train=crime.trn.new[s,] 
crime.test=crime.trn.new[-s,]
```

Here's the glimpse of our training and test datasets for model building & validation:

###Training dataset
```{r}
glimpse(crime.train)
```

###Test dataset
```{r}
glimpse(crime.test)
```


## Build Models
#family=binomial in the glm() function.
```{r}
crime.train.model <- glm(target ~.,family=binomial(link='logit'),data=crime.train)
```
The we obtain the summary of our model
```{r}
summary(crime.train.model)
```
we can see that zn,indus,age,dis and lstat are not statistically significant.
As for the statistically significant variables, rad & nox have  a strong positive association of crime rate while tax has a negative coefficient, suggests as all other variables being equal as tax increases crime rate decreases.

A unit increase in index of accessibility to radial highways increses the log odds by 0.62. Also unit increase in nitrogen oxides concentration increases the logodds by 38.15, while increase in tax rate reduces the log odds by 0.008.

#Now we can run the anova() function on the model to analyze the table of deviance
```{r}
 
anova(crime.train.model, test="Chisq")
```
The difference between the null deviance and the residual deviance shows how our model is doing against the null model (a model with only the intercept). The wider this gap, the better.
Nitrogen oxides concentration is the least deviation, so this variable can be dropped from the model.

```{r}
crime.train.model2 <- glm(target ~ rad + tax ,family=binomial(link='logit'),data=crime.train)
summary(crime.train.model2)
```



## Select Models

## Predictions

## Appendix