---
title: "Critical Thinking Group 4 - HW3"
author: "Sreejaya, Suman, Vuthy"
date: "October 10, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=FALSE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
```

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(car)
library(recommenderlab)
library(knitr)
library(PerformanceAnalytics)
```

## Overview

Dataset  
[Crime - Training data](https://github.com/Nguyver/DATA621-HW/blob/master/HW3/crime-training-data.csv)  
[Crime - Evaluation Data](https://github.com/Nguyver/DATA621-HW/blob/master/HW3/crime-evaluation-data.csv)

## Data Exploration

There are 466 observations, and 14 variables in the given training dataset. The below is a glimpse of the data.
```{r}
crime.trn  <- read.csv("https://raw.githubusercontent.com/Nguyver/DATA621-HW/master/HW3/crime-training-data.csv", header=TRUE, sep=",", stringsAsFactors = FALSE)

crime.evl  <- read.csv("https://raw.githubusercontent.com/Nguyver/DATA621-HW/master/HW3/crime-evaluation-data.csv", header=TRUE, sep=",", stringsAsFactors = FALSE)

glimpse(crime.trn)
```

Lets check the summary of the given dataset, as well as check for any NA values in the data set.
```{r}
summary(crime.trn)
```


There appears to be no missing values. Lets plot the correlation between the variables.

```{r, warning=FALSE, message=FALSE}
cor.matrix <- cor(crime.trn[,1:ncol(crime.trn)])
chart.Correlation(cor.matrix, histogram=TRUE, pch=19)
```

From the above correlation matrix , the target varaible seems to have correlation with 

  * zn
  * indus
  * nox
  * age
  * dis
  * rad
  * tax
  * lstat


## Data Preparation

## Build Models

## Select Models

## Predictions

## Appendix